---
title: "051818"
author: "CC"
date: "May 18, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(BiocInstaller)
```

obtaining the data 
```{r}
counts <- read.csv("data/airway_scaledcounts.csv", stringsAsFactors = FALSE)
metadata <-  read.csv("data/airway_metadata.csv", stringsAsFactors = FALSE)
```

```{r}
colnames(counts)[-1] ==
metadata$id
```

Calculating the mean for the control values

```{r}
#Find id for those labeled controls
control <- metadata[metadata[,"dex"]=="control",]

#Get the mean of the control 
#selecting the 4 ids from our new table "control" to be looked at the table counts. Then we are adding each row and dividing it by the sample number, which is four.
control.mean <- rowSums( counts[ ,control$id] )/ nrow(control)

#If we leave the 4 and our sample number changes then this formula will no longer be calculating the mean. So i changed the 4 for nrow(control)

names(control.mean) <- counts$ensgene
```



Calculating the mean for the treated samples
```{r}
#Find id for those labeled controls
treated <- metadata[metadata[,"dex"]=="treated",]

#Get the mean of the control 
#selecting the 4 ids from our new table "control" to be looked at the table counts. Then we are adding each row and dividing it by the sample number, which is four.
treated.mean <- rowSums( counts[ ,treated$id] )/ nrow(treated)

#If we leave the 4 and our sample number changes then this formula will no longer be calculating the mean. So i changed the 4 for nrow(control)

names(treated.mean) <- counts$ensgene
```


Take a look at each
```{r}
meancounts <- data.frame(control.mean, treated.mean)

colSums(meancounts)
```


Plot it
```{r}
plot(meancounts)
```


Plot with log axis so we can see our data
```{r}
plot(meancounts, log="xy")
```





#Calculate log2foldchange and add it to meancounts
```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)


# NaN (“not a number””) and -Inf (negative infinity) results
```


#The which will tell you what the zero values are true of the meancounts
```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])

head(zero.vals)
```

Now remove the zero count containing genes
```{r}
mycounts <- meancounts[-to.rm,]
head(mycounts)

```

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
paste("up:", sum(up.ind))
paste("down:", sum(down.ind))
```


```{r}
anno <- read.csv("data/annotables_grch38.csv")
head(anno)
head(mycounts)
```


```{r}
results <- merge(mycounts, anno, by.x = "row.names", by.y = "ensgene")
head(results)
```


```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```


To get a list of all available key types

```{r}
columns(org.Hs.eg.db)
```


```{r}
mycounts$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")


#entrez
mycounts$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")


mycounts$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(mycounts),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")

```



```{r}
head(mycounts[up.ind,])
```




#DESeq2 Analysis

Set up DESeq DataSet
```{r}
library(DESeq2)

dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex, 
                              tidy=TRUE)
dds

```


```{r}
dds <- DESeq(dds)
```

```{r}
res <- results(dds)
res
```
We can summarize results
```{r}
summary(res)
```

order my results by p-value
```{r}

resOrdered <- res[order(res$pvalue),]
res05 <- results(dds, alpha=0.05)
summary(res05)
```

Access the actual subset of the data.frame passing a threshold with the subset function

```{r}
resSig05 <- subset(as.data.frame(res), padj < 0.05)
nrow(resSig05)

```

```{r}
resSig01 <- subset(as.data.frame(res), padj < 0.01)
nrow(resSig01)
```

```{r}
resSig01$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="SYMBOL",
                     keytype="ENSEMBL",
                     multiVals="first")

#entrez
resSig01$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")


resSig01$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(resSig01),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")





```






arrange by adjusted p-value
```{r}
ord <- order( resSig01$padj )
#View(res01[ord,])
head(resSig01[ord,])
```

write it up

```{r}
write.csv(resSig01[ord,], "signif01_results.csv")
```


##Plotting
```{r}
i <- grep("CRISPLD2", resSig01$symbol)
resSig01[i,]
```

```{r}
rownames(resSig01[i,])
```

Let's plot the counts with tht gene id. the interesting group variable is the "dex" column

```{r}
plotCounts(dds, gene="ENSG00000103196", intgroup="dex")
```

```{r}
# Return the data
d <- plotCounts(dds, gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
head(d)
```

With this daata we can do a box plot
```{r}
boxplot(count ~ dex , data=d)
```


```{r}
library(ggplot2)
ggplot(d, aes(dex, count)) + geom_boxplot(aes(fill=dex)) + scale_y_log10() + ggtitle("CRISPLD2")
```

