---
title: "Bioinformatics 042518"
author: "CC"
date: "April 25, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Functions again

We can source any file of R code with the 'source'() function

```{r}
source("http://tinyurl.com/rescale-R")
```

Let's make sure things are here

```{r}
ls()
```

Check okur 'rescale()' is working

```{r}
rescale(1:10)
```

```{r, eval=FALSE}
rescale(c(1:10, "string"))
```


Let's check if 'rescale2()' does any better

```{r, eval=FALSE}
rescale2(c(1:10, "string"))
```



##Function for finding missing values in two datasets

Write a 'both_na()' function to do this...

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

is.na(x)
which(is.na(x))
```


##The exclamation mark will give the inverse results
```{r}
which(!is.na(x))

```

##How many total values are we missing
```{r}
sum(is.na(x))

```


```{r}
is.na(x)
is.na(y)
```

##LIke in biology the use of AND
```{r}
is.na(x) & is.na(y)
```


How many TRUE things in total for both x and y in this vector
```{r}
sum( is.na(x) & is.na(y) )
```

My first function can start from this snippet

```{r}
jboth_na <- function(x, y) {
  sum( is.na(x) & is.na(y))
}
```


Test it
```{r}
both_na(x, y)
```


Testing and failing

```{r}
x <- c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)

both_na(x, y2)
```


```{r}
#both_na2( x, y2 )
```



```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3, 4)

ans <- both_na3(x, y)
```


```{r}
ans$which
```

##Last function that is actually useful

```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```


Try the 'intersect()' function and the %in% 
```{r}
intersect(x, y)
x %in% y
```


We can use the logical output in %in%  

```{r}
x [x %in% y]
y [y %in% y]
```

lets put these together as columns of a matrix

```{r}
cbind(x[x %in% y], y[y %in% x])

```

Make it into a first function

```{r}
gene_intersect <- function(x, y) {
cbind( x[ x %in% y ], y[ y %in% x ] )
}
```


Test it
```{r}
gene_intersect(x, y)
```


Lets try with data.frame input rather than vectors

```{r}
gene_intersect2(df1, df2)
```

```{r}
gene_intersect3(df1, df2)
```

Use the 'merge()' function for this

```{r}
merge(df1, df2, by="IDs")
```

```{r}
gene_intersect4(df1, df2)
```


